### 2017-02-05 local-capability-machine.tex
Fixed the writeCond problem using a new notion of "address-stratified" regions.

@@

\begin{proof}[Proof of Lemma~\ref{lem:correctness-g1} (using \texttt{scall} lemma)]
  Let $n$ be given and make the assumptions of the lemma. Define
  \begin{align*}
    W = & [0 \mapsto \iota_{\malloc,0}]\\
        & [1 \mapsto \iota^{\sta,u} (\perma,\ms_\link)]\\
        & [2 \mapsto \iota^\pwl_{\start_\stk, \addrend_\stk}]\\
        & [3 \mapsto \iota^{\nwl,p}_{\start_\adv,\addrend_\adv}]\\
        & [4 \mapsto \iota^\sta (\perma,\ms_{g1} \uplus \ms_\flag)]
  \end{align*}
  and
  \[
    \ms = \ms_{g1} \uplus 
          \ms_\flag \uplus                
          \ms_\link \uplus 
          \ms_\adv \uplus 
          \ms_\malloc 
  \]
  If we can show
  \begin{equation}
    \label{eq:f4:start-conf-in-obs}
        \npair{(\reg_0\update{\pcreg}{c_\adv}\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}},\ms \uplus \ms_\stk)} \in \observations(W)
  \end{equation}
  then, the termination assumption gives us that part of $m$ satisfies a private future world of $W$. Region 4 is permanent, so
  \[
    m(\flag) = 0
  \]
  So it suffices to show \ref{eq:f4:start-conf-in-obs}. 
  To this end use the FTLR to show $\npair{c_\adv} \in \stder(W)$, so show
  \begin{enumproof}
    \item $\npair{(\start_\adv,\addrend_\adv)} \in \readCond{}(\glob)(W)$\\
      Simple using region $3$ in $W$ and Lemma~\ref{lem:nwlp-subset-pwl}.
    \item $\npair{(\start_\adv,\addrend_\adv)} \in \writeCond{}(\iota^\nwl,\glob)(W)$\\
      Simple using region $3$ in $W$.
  \end{enumproof}
  in conclusion $\npair{c_\adv} \in \stder(W)$. We get \ref{eq:f4:start-conf-in-obs} if we show the following
  \begin{enumproof}[resume]
    \item $\memSat{\ms \uplus \ms_\stk}{W}$\\
      \begin{enumproof}
        \item $\memSat{\ms_{g1} \uplus \ms_\flag}{[4 \mapsto \iota^\sta (\perma,\ms_{g1} \uplus \ms_\flag)]}$ \\
          Lemma~\ref{lem:mem-sat-static}.
        \item $\memSat{\ms_\stk}{[2 \mapsto \iota^\pwl_{\start_\stk,\addrend_\stk}]}$\\
          Lemma~\ref{lem:mem-sat-data-only-std-regions} and assumption that $\ms_\stk$ is all 0.
        \item $\memSat{\ms_{\malloc} \uplus \ms_\link \uplus \ms_\adv}{[0 \mapsto \iota_{\malloc,0}][1 \mapsto \iota^{\sta,u} (\perma,\ms_\link)][3 \mapsto \iota^{\nwl,p}_{\start_\adv,\addrend_\adv}]}$\\
          For convenience define
          \[
            W_{\var{mini}} = [0 \mapsto \iota_{\malloc,0}][1 \mapsto \iota^{\sta,u} (\perma,\ms_\link)][3 \mapsto \iota^{\nwl,p}_{\start_\adv,\addrend_\adv}]
          \]
          Partitioning the memory segment in the components of the disjoint union, the $\malloc$ part follows from assumption $\memSat{\ms_\malloc}{[0 \mapsto \iota_{\malloc,0}]}$ and the $\malloc$ specification.

@@

\subsection{The logical relation}
The logical relation is defined by several mutual recursive definitions. In order to handle this mutual recursion and show that this definitions are well-defined, Banach's fixed-point theorem can be used. % TODO: Reference?
We have omitted the details of this construction here, but it is done by parameterising all the definitions by the value relation.
\begin{align*}
  & \writeCond{} : (\Regions \times \Globals) \fun \Worlds \monnefun \UPred{\Addrs^2}  \\
  & \writeCond{}(\iota,\gl)(W) =  \\
  & \quad \begin{aligned}[t]
    \{ \npair{(\start,\addrend)} \mid & \;\exists r \in \var{localityReg}(g,W) \ldotp \\
    & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
    & \;\qquad W(r)\nsupsim[n-1] \iota_{\start',\addrend'} \}
  \end{aligned}
\end{align*}

@@

@@

@@

\begin{lemma}
  \label{lem:revoketemp-writecond}
  If
  \[
    \npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(\revokeTemp{W})
  \]
  then
  \[
    \npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(W)
  \]
\end{lemma}
\begin{proof}[Proof of Lemma~\ref{lem:revoketemp-writecond}]
  \[
    \npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(\revokeTemp{W})
  \]
  Gives $r \in \localityReg(\gl,\revokeTemp{W})$ such that
  \[
    \forall [\start',\addrend'] \subseteq [\start,\addrend] \ldotp \revokeTemp{W}(r) \nsupsim[n-1] \iota_{[\start',\addrend']}
  \]
  Notice $\revokeTemp{W}(r)$ is a $\perma$ region, so $\revokeTemp{W}(r) = W(r)$. Using $r$ as witness, the result is immediate.
\end{proof}

@@

 \begin{lemma}[pwl writecond implies nwl]
   \label{lem:pwl-writecond-implies-nwl}
   If $\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W)$ then
   $\npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \}$.
 \end{lemma}
 \begin{proof}
   \begin{enumproof}
   \item $\exists r \in \var{localityReg}(g,W) \ldotp$ $\exists
     [\start',\addrend'] \supseteq [\start,\addrend] \ldotp$ $W(r)\nsupsim[n-1]
     \iota^\pwl_{\start',\addrend'}$: by definition of $\writeCond{}$.
   \item Suffices: $W(r)\nsupsim[n-1] \iota^\nwl_{\start',\addrend'}$. By
     definition of $\writeCond{}$
   \item $W(r)\nsupsim[n-1] \iota^\pwl_{\start',\addrend'} \nsupsim[n-1]
     \iota^\nwl_{\start',\addrend'}$: follows by Lemma~\ref{lem:nwl-subset-pwl}.
   \end{enumproof}
 \end{proof}

@@

   then $\addr \in \dom(\ms')$ (i.e. $\ms\update{a}{w} =
   \ms'\update{a}{w}\uplus\ms_f$) and
   $\heapSat[{\ms'\update{\addr}{\var{w}}}]{n}{W}$
 \end{lemma}
 \begin{proof}
   TODO
 \end{proof}

@@

\begin{lemma}[Stack capability in value relation]
  \label{lem:stack-cap-vr}
  If
  \begin{itemize}
  \item \pointstostack{\reg}{\emptyset}{\ms}
  \item $\exists r \ldotp W(r) = \iota^\pwl (\dom(\ms))$
  \end{itemize}
  then
  \[
    \npair{\reg(r_\stk)} \in \stdvr(W)
  \]
\end{lemma}
\begin{proof}[Proof of Lemma~\ref{lem:stack-cap-vr}]
  Say 
\[
  \reg(r_\stk) = c_\stk = ((\rwlx,\local),\start,\addrend,\_)
\]
  Show
  \begin{enumproof}
    \item $\npair{(\start,\addrend)} \in \readCond{}(\local)(W)$ : \label{pf:stack-cap-read}\\
      Amounts to
      \[
        \iota^\pwl (\dom(\ms)) \nsubsim \iota^\pwl_{\start,\addrend}
      \]
      which is true as they are even equal.
    \item $\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\local)(W)$ : \label{pf:stack-cap-write}\\
      Amounts to
      \[
        \iota^\pwl (\dom(\ms)) \nsupsim \iota^\pwl_{\start,\addrend}
      \]
      which is true as they are even equal.
      \item $\npair{(\rwlx,\start_\stk, \addrend_\stk)} \in \execCond{}(\local)(W)$\\
        Using \ref{pf:stack-cap-write} and \ref{pf:stack-cap-read}, we can use Lemma~\ref{lem:stack-exec-cond-help}.
      \item $\npair{(\rwx,\start_\stk, \addrend_\stk)} \in \execCond{}(\local)(W)$\\
        Using \ref{pf:stack-cap-write} and \ref{pf:stack-cap-read}, we can use Lemma~\ref{lem:stack-exec-cond-help}.
      \item $\npair{(\exec,\start_\stk, \addrend_\stk)} \in \execCond{}(\local)(W)$\\
        Using \ref{pf:stack-cap-read} and \ref{pf:stack-cap-write}, we can use Lemma~\ref{lem:stack-exec-cond-help}.
  \end{enumproof}
\end{proof}

>>>>

        Using Lemma~\ref{lem:iota-pwl-address-stratified}, this amounts to showing

@@

\begin{proof}[Proof of Lemma~\ref{lem:correctness-g1} (using \texttt{scall} lemma)]
  Let $n$ be given and make the assumptions of the lemma. Define
  \begin{align*}
    W = & [0 \mapsto \iota_{\malloc,0}]\\
        & [1 \mapsto \iota^{\sta,u} (\perma,\ms_\link)]\\
        & [2 \mapsto \iota^\pwl_{\start_\stk, \addrend_\stk}]\\
        & [3 \mapsto \iota^{\nwl,p}_{\start_\adv,\addrend_\adv}]\\
        & [4 \mapsto \iota^\sta (\perma,\ms_{g1} \uplus \ms_\flag)]
  \end{align*}
  and
  \[
    \ms = \ms_{g1} \uplus 
          \ms_\flag \uplus                
          \ms_\link \uplus 
          \ms_\adv \uplus 
          \ms_\malloc 
  \]
  If we can show
  \begin{equation}
    \label{eq:f4:start-conf-in-obs}
        \npair{(\reg_0\update{\pcreg}{c_\adv}\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}},\ms \uplus \ms_\stk)} \in \observations(W)
  \end{equation}
  then, the termination assumption gives us that part of $m$ satisfies a private future world of $W$. Region 4 is permanent, so
  \[
    m(\flag) = 0
  \]
  So it suffices to show \ref{eq:f4:start-conf-in-obs}. 
  To this end use the FTLR to show $\npair{c_\adv} \in \stder(W)$, so show
  \begin{enumproof}
    \item $\npair{(\start_\adv,\addrend_\adv)} \in \readCond{}(\glob)(W)$\\
      Simple using region $3$ in $W$ and Lemma~\ref{lem:nwlp-subset-pwl}.
    \item $\npair{(\start_\adv,\addrend_\adv)} \in \writeCond{}(\iota^\nwl,\glob)(W)$\\
      Simple using region $3$ in $W$, using
      Lemma~\ref{lem:iota-nwl-address-stratified}.
  \end{enumproof}
  in conclusion $\npair{c_\adv} \in \stder(W)$. We get \ref{eq:f4:start-conf-in-obs} if we show the following
  \begin{enumproof}[resume]
    \item $\memSat{\ms \uplus \ms_\stk}{W}$\\
      \begin{enumproof}
        \item $\memSat{\ms_{g1} \uplus \ms_\flag}{[4 \mapsto \iota^\sta (\perma,\ms_{g1} \uplus \ms_\flag)]}$ \\
          Lemma~\ref{lem:mem-sat-static}.
        \item $\memSat{\ms_\stk}{[2 \mapsto \iota^\pwl_{\start_\stk,\addrend_\stk}]}$\\
          Lemma~\ref{lem:mem-sat-data-only-std-regions} and assumption that $\ms_\stk$ is all 0.
        \item $\memSat{\ms_{\malloc} \uplus \ms_\link \uplus \ms_\adv}{[0 \mapsto \iota_{\malloc,0}][1 \mapsto \iota^{\sta,u} (\perma,\ms_\link)][3 \mapsto \iota^{\nwl,p}_{\start_\adv,\addrend_\adv}]}$\\
          For convenience define
          \[
            W_{\var{mini}} = [0 \mapsto \iota_{\malloc,0}][1 \mapsto \iota^{\sta,u} (\perma,\ms_\link)][3 \mapsto \iota^{\nwl,p}_{\start_\adv,\addrend_\adv}]
          \]
          Partitioning the memory segment in the components of the disjoint union, the $\malloc$ part follows from assumption $\memSat{\ms_\malloc}{[0 \mapsto \iota_{\malloc,0}]}$ and the $\malloc$ specification.

@@

\subsection{The logical relation}
The logical relation is defined by several mutual recursive definitions. In order to handle this mutual recursion and show that this definitions are well-defined, Banach's fixed-point theorem can be used. % TODO: Reference?
We have omitted the details of this construction here, but it is done by parameterising all the definitions by the value relation.

\begin{align*}
&\iota = (v,s,\phi_\pub,\phi,H) \text{ is address-stratified iff }\\
&\qquad\begin{multlined}
  \forall s', \hat{W}, n, \ms, \ms' \ldotp \\
  \npair{\ms},\npair{\ms'} \in H~ s'~ \hat{W} \Rightarrow \\
  \dom(\ms) = \dom(\ms') \wedge \\
  \forall \addr \in
  \dom(\ms)\ldotp \npair{\ms\update{\addr}{\ms'(\addr)}} \in H~ s'~ \hat{W}
\end{multlined}
\end{align*}

\begin{align*}
  & \writeCond{} : (\Regions \times \Globals) \fun \Worlds \monnefun \UPred{\Addrs^2}  \\
  & \writeCond{}(\iota,\gl)(W) =  \\
  & \quad \begin{aligned}[t]
    \{ \npair{(\start,\addrend)} \mid & \;\exists r \in \var{localityReg}(g,W) \ldotp \\
    & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
    & \;\qquad W(r)\nsupsim[n-1] \iota_{\start',\addrend'} \text{ and }\\
    & \;\qquad W(r) \text{ is address-stratified } \}
  \end{aligned}
\end{align*}

@@

\begin{lemma}
  \label{lem:iota-pwl-address-stratified}
  $\iota^\pwl_{\start,\addrend}$ is address-stratified.
\end{lemma}
\begin{proof}
  Easy unfolding of definitions.
\end{proof}

@@

\begin{lemma}
  \label{lem:iota-nwl-address-stratified}
  $\iota^\nwl_{\start,\addrend}$ is address-stratified.
\end{lemma}
\begin{proof}
  Easy unfolding of definitions.
\end{proof}

@@

\begin{lemma}
  \label{lem:revoketemp-writecond}
  If
  \[
    \npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(\revokeTemp{W})
  \]
  then
  \[
    \npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(W)
  \]
\end{lemma}
\begin{proof}[Proof of Lemma~\ref{lem:revoketemp-writecond}]
  \[
    \npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(\revokeTemp{W})
  \]
  Gives $r \in \localityReg(\gl,\revokeTemp{W})$ such that
  \[
    \forall [\start',\addrend'] \subseteq [\start,\addrend] \ldotp \revokeTemp{W}(r) \nsupsim[n-1] \iota_{[\start',\addrend']}
  \]
  and 
  \begin{equation*}
    \revokeTemp(W)(r) \text{ is address-stratified}
  \end{equation*}
  Notice $\revokeTemp{W}(r)$ is a $\perma$ region, so $\revokeTemp{W}(r) = W(r)$. Using $r$ as witness, the result is immediate.
\end{proof}

@@

 \begin{lemma}[pwl writecond implies nwl]
   \label{lem:pwl-writecond-implies-nwl}
   If $\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W)$ then
   $\npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \}$.
 \end{lemma}
 \begin{proof}
   \begin{enumproof}
   \item $\exists r \in \var{localityReg}(g,W) \ldotp$ $\exists
     [\start',\addrend'] \supseteq [\start,\addrend] \ldotp$ $W(r)\nsupsim[n-1]
     \iota^\pwl_{\start',\addrend'}$ and $W(r)$ is address-stratified: by definition of $\writeCond{}$.
   \item Suffices: $W(r)\nsupsim[n-1] \iota^\nwl_{\start',\addrend'}$. By
     definition of $\writeCond{}$
   \item $W(r)\nsupsim[n-1] \iota^\pwl_{\start',\addrend'} \nsupsim[n-1]
     \iota^\nwl_{\start',\addrend'}$: follows by Lemma~\ref{lem:nwl-subset-pwl}.
   \end{enumproof}
 \end{proof}

@@

   then $\addr \in \dom(\ms')$ (i.e. $\ms\update{a}{w} =
   \ms'\update{a}{w}\uplus\ms_f$) and
   $\heapSat[{\ms'\update{\addr}{\var{w}}}]{n}{W}$
 \end{lemma}
 \begin{proof}
   \begin{enumproof}
   \item $\npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(W)$ where $\iota
     = \iota^\pwl$ or $\iota = \iota^\nwl$ and (if $\var{w} =
     ((\_,\local),\_,\_,\_)$, then $\iota = \iota^\pwl$).

     By definition of $\stdvr(W)$ and $\writeAllowed{}$, from
     $\npair{c}\in\stdvr(W)$, $((\perm,\gl),\start,\addrend,\addr) = c$ and
     $\writeAllowed{\perm}$ and the fact that (if $\var{w} =
     ((\_,\local),\_,\_,\_)$, then $\perm \in \{\rwlx,\readwritel \}$)
   \item $\exists r \in \var{localityReg}(\gl,W) \ldotp$ $\exists
     [\start',\addrend'] \supseteq [\start,\addrend] \ldotp$ $W(r)\nsupsim[n-1]
     \iota_{\start',\addrend'}$ and $W(r)$ is address-stratified. By definition
     of $\writeCond{}$.
   \item $\exists P : \activeReg{W} \rightarrow \HeapSegments \ldotp$
     $\memSatPar{\ms'}{W}{P}$. By definition of $\heapSat[\ms']{n}{W}$.
   \item $\ms' = \biguplus_{r\in\activeReg{W}}P(r)$ and $\forall r \in
     \activeReg{W} \ldotp$ $\exists H,s \ldotp$ $W(r) = (\_,s,\_,\_,H)$ and
     $\npair[n]{P(r)} \in H(s)(\xi^{-1}(W))$. By definition of
     $\memSatPar{\ms'}{W}{P}$.
   \item $\exists H,s \ldotp$ $W(r) = (\_,s,\_,\_,H)$ and
     $\npair[n]{P(r)} \in H(s)(\xi^{-1}(W))$. By instantiating the previous
     point to the $r$ from the $\writeCond{}$.
   \item $\npair{w} \in \iota.H~(\iota.s)~(\xi^{-1}(W))$ by definition of
     $\iota^\pwl$, $\iota^\nwl$ and the fact that (if $\var{w} =
     ((\_,\local),\_,\_,\_)$, then $\iota = \iota^\pwl$).
   \item Define $\ms'_w$ such that $\dom(\ms'_w) = [\start',\addrend']$,
     $\ms'_w(\addr) = w$ and $\ms'_w(\addr') = 0$ for $\addr' \neq \addr$. It's
     easy to show from the previous point that $\npair{\ms'_w} \in
     H(s)(\xi^{-1}(W))$.
   \item $\dom(P(r)) = \dom(\ms'_w) = [\start',\addrend'] \ni \addr$ and
     $\npair{P(r)\update{\addr}{w}} \in H(s)(\xi^{-1}(W))$ by applying the fact
     that $W(r)$ is address-stratified, combined with the previous point.
   \item Define $P'(r) = P(r)\update{a}{w}$ and $P'(r') = P(r')$ for $r' \neq r$.
   \item $\ms'\update{a}{w} = \biguplus_{r\in\activeReg{W}}
     P'(r)$ and $\ms'\update{a}{w} :_{n,P'} W$. By definition of
     $\memSatPar{\ms'}{W}{P}$ and the previous two points.
   \end{enumproof}
 \end{proof}

@@

\begin{lemma}[Stack capability in value relation]
  \label{lem:stack-cap-vr}
  If
  \begin{itemize}
  \item \pointstostack{\reg}{\emptyset}{\ms}
  \item $\exists r \ldotp W(r) = \iota^\pwl (\dom(\ms))$
  \end{itemize}
  then
  \[
    \npair{\reg(r_\stk)} \in \stdvr(W)
  \]
\end{lemma}
\begin{proof}[Proof of Lemma~\ref{lem:stack-cap-vr}]
  Say 
\[
  \reg(r_\stk) = c_\stk = ((\rwlx,\local),\start,\addrend,\_)
\]
  Show
  \begin{enumproof}
    \item $\npair{(\start,\addrend)} \in \readCond{}(\local)(W)$ : \label{pf:stack-cap-read}\\
      Amounts to
      \[
        \iota^\pwl (\dom(\ms)) \nsubsim \iota^\pwl_{\start,\addrend}
      \]
      which is true as they are even equal.
    \item $\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\local)(W)$ : \label{pf:stack-cap-write}\\
      Using Lemma~\ref{lem:iota-pwl-address-stratified}, this amounts to
      \[
        \iota^\pwl (\dom(\ms)) \nsupsim \iota^\pwl_{\start,\addrend}
      \]
      which is true as they are even equal.
      \item $\npair{(\rwlx,\start_\stk, \addrend_\stk)} \in \execCond{}(\local)(W)$\\
        Using \ref{pf:stack-cap-write} and \ref{pf:stack-cap-read}, we can use Lemma~\ref{lem:stack-exec-cond-help}.
      \item $\npair{(\rwx,\start_\stk, \addrend_\stk)} \in \execCond{}(\local)(W)$\\
        Using \ref{pf:stack-cap-write} and \ref{pf:stack-cap-read}, we can use Lemma~\ref{lem:stack-exec-cond-help}.
      \item $\npair{(\exec,\start_\stk, \addrend_\stk)} \in \execCond{}(\local)(W)$\\
        Using \ref{pf:stack-cap-read} and \ref{pf:stack-cap-write}, we can use Lemma~\ref{lem:stack-exec-cond-help}.
  \end{enumproof}
\end{proof}
